The five pillars of computational reproducibility

### 2023-05-15 docs/5pillars.Rmd
progress

A number of guides recommending enhanced computational reproducibility practices have emerged
[@Peng2011-ia; @Tan2010-is; @Sandve2013-lk; @Piccolo2016-ji; @Gruning2018-sg; @Lewis2016-qh;
@Brito2020-ld; @Rodrigues2023-fw; @Wilson2017-uq; @Akalin2018-eg ]
and this article attempts to unify these concepts and assemble a group of best practices where
computational reproducibility can be considered as complete, using a framework called the
five pillars of reproducible computational research (**Figure 1**).

The intended audience is bioinformatics data analysts (not tool developers), however the
principles described here should translate to other domains of study.

@@

Getting started with R and python:

@@

Practical guides for jupyter and R Markdown - put these on webarchive and provide persistent links.

@@

Practical guides for `git` and GitHub - put these on webarchive and provide persistent links.

@@

@@

@@

RESEARCHOPS: A. PRINCIPLED FRAMEWORK. AND GUIDE TO. COMPUTATIONAL. REPRODUCIBILITY.
https://osf.io/ubj76/download

@@

@@

@@

>>>>

A number of guides recommending enhanced computational reproducibility practices have emerged
[@Peng2011-ia; @Tan2010-is; @Sandve2013-lk; @Piccolo2016-ji; @Gruning2018-sg; @Lewis2016-qh;
@Brito2020-ld; @Rodrigues2023-fw; @Wilson2017-uq; @Akalin2018-eg ]
and this article attempts to unify these concepts and assemble a group of best practices where
computational reproducibility can be considered as complete, using a framework called the
five pillars of reproducible computational research (**Figure 1**).
The intended audience is bioinformatics data analysts (not tool developers), however the
principles described here should translate to other domains of study.

@@

@@

@@

@@

Use webarchive to archive important resources.

@@

These innovations, when brought together facilitate extreme reproducibility, ie: makes reproducing a previous study relatively
easy in terms of complexity, number of commands, and in time.
For example, we made an example transcriptome analysis workflow based on the five pillars approach and it could could be installed
and generate differential expression and pathway enrichment results in a matter of minutes.

With the above innovations, it's possible to reproduce a computational study with just a few

## References to read

@@

RESEARCHOPS: A PRINCIPLED FRAMEWORK AND GUIDE TO COMPUTATIONAL REPRODUCIBILITY.
https://osf.io/ubj76/download

@@

## Practical guides

@@

Put these on webarchive and provide persistent links

@@

### Getting started with data analysis in R and python

### Literate programming with jupyter, R Markdown and Quarto

### Practical guides for `git` and GitHub

### Practical guides for Conda, Guix and Docker

### Practical guides for documenting computational research

### Practical guides to data sharing
